Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-07-02 kappa.tex
Add more references in the introduction.

% ice volume 26.07 + 2.93 = 29.00
% meters sea-level 58.3 + 0.70 = 59.0
Together, the Greenland and Antarctic ice sheets currently host most glacial
ice on the planet, corresponding to 59~meters of potential sea-level rise
\citep{Bamber.etal.2001, Fretwell.etal.2013}. At the apogee of the last glacial
cycle, the last glacial maximum, sea-level was 120 to 135\,m lower than today
\citep{Clark.Mix.2002}, implying an additional, double amount of ice stored in
ice sheets and glaciers on land. Systematic mapping and compilation of glacial
landforms from all continents permitted reconstructions of this former ice
cover \citep{Ehlers.Gibbard.2007}, indicating that ice volume
was repartitioned between more extensive Greenland and Antarctic ice sheets,
the Scandinavian and Barents-Kara ice sheets in the Eurasia arctic, and the
Innuitian, Laurentide and Cordilleran ice sheets over North America
(\cref{fig:paleo-glaciation}).

Nevertheless, the landscape record is timely sparse and spatially incomplete,
and geological reconstructions of former glacier extent demand a high amount of
interpretation. Palaeo-ice sheets did not leave a systematic record of their
passage, and much of the evidence has been overridden by multiple glacier
re-advance and other geomorphological processes. This is where numerical
models, incorporating the processes of glacier sliding and ice deformation
observed on modern glaciers, can help. Although these models alone contain too
many uncertainties to reproduce accurate picture of past glacial cycles, they
can be calibrated against field-based evidence where it exists, while allowing
a physics-based, quantitative reconstruction where it does not.

>>>>

% ice volume 26.07 + 2.93 = 29.00
% meters sea-level 58.3 + 0.70 = 59.0
Together, the Greenland and Antarctic ice sheets currently host most glacial
ice on the planet, corresponding to 59~meters of potential sea-level rise
\citep{Bamber.etal.2001, Fretwell.etal.2013}. At the apogee of the last glacial
cycle, the last glacial maximum, sea-level was 120 to 135\,m lower than today
\citep{Clark.Mix.2002}, implying an additional, double amount of ice stored in
ice sheets and glaciers on land. Systematic mapping and compilation of glacial
landforms from all continents permitted reconstructions of this former ice
cover \citep{Ehlers.Gibbard.2007}, indicating that ice volume
was repartitioned between more extensive Greenland and Antarctic ice sheets,
the Scandinavian and Barents-Kara ice sheets in the Eurasian arctic, and the
Innuitian, Laurentide and Cordilleran ice sheets over North America
(\cref{fig:paleo-glaciation}).

Nevertheless, the landscape record is sparse in time and most often spatially
incomplete. Palaeo-ice sheets did not leave a systematic record of their
passage, and much of the evidence has been overridden by subsequent glacier
re-advances and other geomorphological processes \citep{Kleman.1994}.
Therefore, geomorphological reconstructions of former glacier extent require a
substantial amount of interpretation. This is where numerical models,
incorporating the processes of glacier sliding \citep[e.g.,][]{Weertman.1957}
and deformation \citep[e.g.,][]{Nye.1953} observed on modern glaciers, can help.
Although these models alone contain too many uncertainties to reproduce
accurate picture of past glacial cycles \citep[e.g.,][]{Hebeler.etal.2008},
they can be calibrated against geological data where they exist, while
allowing a physics-based, quantitative reconstruction where they do not.
